YALE DEPARTMENT OF ECONOMICS \\ \vspace*{60pt} The Folly of the Crowd: \\ Market Inefficiencies in Peer-to-Peer Lending

### 2015-03-31 draft.Rnw
Final Draft. Changed abstract.

\title{YALE DEPARTMENT OF ECONOMICS \\ \vspace*{60pt} The Folly of the Crowd: \\
Market Inefficiencies in Peer-to-Peer Lending}
\author{Samuel Shleifer
  \footnote{Thanks to Professor Vineet Kumar for his help
  in developing the ideas and statistical methodologies involved in this paper,
  to Devin Pope and Rajkamal Iyer for help manipulating the Prosper data, 
  and to Andrei Shleifer, Nicholas Barberis and Andrew Metrick for helpful comments and suggestions.}
  \footnote{R code available at \url{github.com/sshleifer/ProsperThesis}}\\
Thesis Advisor: Vineet Kumar\footnote{Yale School of Management}}
\maketitle

@@

\afterpage{\blankpage}
\newpage
\begin{abstract}
This essay analyzes the effect of market maker interventions on financial market
efficiency, using data from Prosper.com, the first p2p lending
website in the US.\footnote{P2P lending is the practice of individual and
  institutional investors lending money to borrowers over the internet, without
the presence of a financial intermediary} Between 2006 and 2011, Prosper underwent two
major policy changes that guided and then forced investors to lend at higher
interest rates, as well as to reduce their exposure to subprime borrowers.

@@

This essay expands on the work of Freedman and Jin (2008), who use investor
learning to explain the improvement, with the benefit of three more years of
data and two major policy changes. 
Instead of investor learning as the catalyst that led to a more efficient market, I frame Prosper's rule changes as catalysts that guide
and then force investors towards making better decisions, and find that when
investors set rates by auction, investment strategies using publicly available
information can earn excess returns, an inconsistency with the predictions of
the Efficient Markets Hypothesis, but that once Prosper begins setting rates,
it is no longer possible to beat the market.
\end{abstract}
\newpage
\tableofcontents

@@

In the P2P Lending context, the market's price is the interest rate charged on a loan. This paper studies the question of whether interest rates on Prosper.com are set efficiently. (Fama, 1970)'s semi-strong EMH predicts that if interest rates were set efficiently, then...
\begin{enumerate}
\item Loans would have positive expected returns. Investors would be unwilling
  to lend to Borrowers if they expected to lose money. \footnote{Assuming a
  Risk-Free Rate of 0.}
\item Riskier loans would have higher expected returns. \footcite{sharpe}
\item It is not possible to do better than the average investor using publicly available information. \footcite{fama}
\end{enumerate}
Hypothesis 1 is tested by measuring historical returns of loans to borrowers of
each credit grade.
Hypothesis 2 is tested by comparing the returns of high risk loans to low risk
loans.
Hypothesis 3 is tested by using publicly available information to select loans
with a higher than average expected return, without look-ahead bias.

Prosper.com data provide some additional valuable insights into investor
behavior. Over the period I study, there are three interest rate determination regimes:
\begin{enumerate}
\item Early Auction Regime (April 19, 2006-October 15, 2008): Rates are set by investor auction. Investors have large amounts of data, but Prosper does not guide them meaningfully.
\item Late Auction Regime: (July 20, 2009-December 19, 2010): Rates are set by investor auction, with Prosper offering valuable guidance in the form of more complex credit grading.
\item Prosper-rate Regime: (December 20, 2010 - August 3, 2011): Prosper sets rates using its more complex credit grading algorithm.
\end{enumerate}
Additionally, for regimes (2) and (3), subprime borrowers are not allowed to
apply for Prosper loans.

@@

\begin{table}
  %\renewcommand\arraystretch{1.4}\arrayrulecolor{LightSteelBlue3}
  %\captionsetup{singlelinecheck=false, font=blue, labelfont=sc, labelsep=quad}
  \caption{Timeline of Events}\vskip -1.5ex
  \begin{tabular}{@{\,}r <{\hskip 2pt} !{\foo}
    >{\raggedright\arraybackslash}p{12cm}}
    \toprule
    \addlinespace[1.5ex]
    Apr. 21, 2006 & First loan of early auction regime.\\
    Mar. 1, 2007 & Prosper bans borrowers with credit score $\leq$ 520.\\
    Oct. 15, 2008 & Early Auction period ends with SEC enforced shutdown.\\
    Jul. 20, 2009 & Prosper reopens with more informative credit grades, requires repeat borrowers to have credit score
    $\geq$ 600, and requires new borrowers to have credit score $\geq$ 640.
    Beginning of Early Auction Regime.\\
    Dec. 19, 2010 & Prosper-rate regime, where Prosper sets rates directly,
    begins, because "auctions made the deployment of funds more time consuming
    with little gain in lender returns,” - Prosper CEO Chris
    Larsen.\footcite{renton}\\
    Aug. 3, 2011 & Last loan in Prosper-rate regime originated.
  \end{tabular}
\label{timeline}
\end{table}

@@

As an illustrative example, suppose Bill the borrower needed to borrow \$100, and set his maximum rate at 20\%.
Let Chris, David and Ellen be investors, with Chris bidding (\$50, 15\%), David bidding (\$50, 18\%) and Ellen bidding (\$30,19\%). In this scenario, Chris and David would win the auction, and each lend lend Bill \$50, and receive an 18\% interest rate. Bill would pay 18\% interest before fees.
\subsection{Early Auction Regime}
In the early auction regime, loans originated between April 19, 2006 and October
15, 2008, auction participants had access to considerable information from
Prosper, including over 60  Borrower characteristics, such as residence,
employment status and an Experian credit score bucket, but few intelligent risk estimates. Additionally, Prosper did not regulate who could apply for a loan on the website. Besides basic fraud prevention, any applicant who received 100\% funding from investors was eligible, for Prosper's first year of operation. On March 1, 2007, amidst poor returns, Prosper announced that listings with credit scores below 520 would no longer be offered to investors. This intervention is not discussed for lack of space, but shows that after a year, more or less, Prosper felt the need to nudge a fully decentralized investment process in the right direction.

@@

Fama, (1970) notes the dependence of most tests of market efficiency on a model of risk and return.
The models, including most famously, Sharpe's capital asset pricing model ('CAPM'), make two predictions:
First,  since an investor can always hold cash risklessly, the average return on risky investments must be positive. Second, securities with higher systematic risk earn higher average returns.\footcite{sharpe}

@@

\subsection{Soft Information and Credit Screening}
The novel idea of P2P is a large group of investors may be better equipped to assess risk than traditional financial institutions. ``Traditionally, the credit score provided by rating agencies has been the main tool banks use to screen smaller borrowers."\footcite{iyer} As such, the initial strand of literature focuses largely on the novel sources of information utilized by P2P investors.
Gao and Lin (2012) analyze the text written by the borrower in the ``description
field" of the application, and find that descriptions that are easy to read have
2.3\% lower default rates than average, while complex narratives are associated
with a 3.6\% higher default rate, but do not analyze return on
investment.\footcite{gao} Pope and Sydnor (2008) find that investors can reduce
their defaults by statistically discriminating against racial minorities, but do
not analyze return on investment. Iyer et. al (2009) find that within the credit score buckets provided by Prosper, investors use soft information to further sort borrowers within buckets, reducing defaults.\footcite{iyer}

@@

\section{Data}
My sample consists of 39,555 three year, unsecured (no collateral) Loans, worth
\$231.7 Million,  originated from 2006-2011. The data was downloaded off Prosper.com.
Borrowers apply for loans by creating a public listing on the Prosper.com website, and can request anywhere from \$1,000 to \$35,000.\footnote{The minimum increased to \$2,000 effective Nov. 1, 2011}.
Since our analysis requires measurement of loan performance, which can only take
place three  years after origination (given 3 year maturity on all loans), the
sample is ended in 2011, as we do not know what will end up happening to more recently originated loans. 
As shown in Table \ref{sample}, this leaves us with many more early auction regime loans than the post shutdown regimes.
<<sampTab, cache=FALSE>>=
samp.tab = mf(count(nd,regime))
colnames(samp.tab) = c('Regime','n','Freq.')
xtable(samp.tab,kdigits=4, caption='Observations per regime. Each observation is
one funded loan.', label='sample')
@

To help investors better assess the creditworthiness of borrowers, for every
loan originated in the site's history, Prosper posts historical data regarding loan
performance, such as payments made, interest rate, and principal lost. Borrower
characteristics, including credit variables, which are usually obtained directly
from Experian and verification of income and employment, are available to
investors for each listing. Moreover, historical market data is available for
download, through a query tool, and through third party websites. 
Interviews conducted at the 2008 Prosper Days Conference suggest that there was
``enormous heterogeneity in lender awareness of the data, ability to process the
data, and intent to track the data over time.''\footcite{freedman}

\subsection{Variable Definitions for the Prosper Data}
<<sumTable,cache=FALSE>>=
nd$default.rate =  nd$defaulted
k4 = c('roi','default.rate', 'LenderYield','BorrowerRate','regime','LoanOriginalAmount','Investors','CreditScoreEst','DebtToIncomeRatio', 'InquiriesLast6Months')
ag4 <- aggregate(.~regime, nd[k4], function(x) c(mean = mean(x), sd = sd(x)))
d4 = t(do.call(data.frame, c(ag4, check.names = FALSE)))
d5 = data.frame(d4)
names(d5) = d4[1,]
d6 = d5[-1,]
xtable(d6,caption='This table presents means (.mean) and standard deviations
(.sd) for relevant variables by regime.',digits=3, label='sumStats')
@
For each loan, Prosper offers 63 variables, I now define an important subset.
\begin{itemize}
\item BorrowerRate: Annual interest rate paid by the borrower
\item TotalCreditLinespast7years: Number of credit lines, including cards and loans, used by the borrower in the seven years before the prosper loan was originated.
\item Inquirieslast6Months: Number of times a lender inquired about the Borrower's credit report, in the 6 months before the loan was originated. Traditional lenders make inquiries when a borrower applies for a loan, so this serves as a proxy for the number of times someone has borrowed in the past 6 months.
\item CurrentDelinquencies: Number of loans on which the borrower is delinquent at the origination date.
\item DelinquenciesLast7Years: The number of loans which the borrower has not repaid fully over the 7 years before the origination date.
\item MonthlyLoanPayment: The borrower's scheduled monthly payment on the Prosper loan.
\item CreditScoreRangeLower: The lower bound of the credit score range (which
  bounds his raw credit score) , that bounds a borrower's credit grade, as
  provided by Experian. Prosper provides the range instead of the raw score to
  preserve borrower anonymity.
\item CreditScoreEst: The average of the lower and upper bounds of the credit score range. 
\item PublicRecordsLast10Years: Number of public records of borrowing activity in the past 10 years at the origination date. 
\item DebtToIncomeRatio: The debt to income ratio of the borrower at the origination date. This value is capped at 10.01 (any debt to income ratio larger than 1000\% will be coded as 1001\%).
\item TotalProsperPaymentsBilled: Number of on time payments the borrower made on Prosper loans at the time they created the listing. This value is 0 if the borrower had no prior loans.
\item LenderYield: The lender's Yield on the loan. Equal to the Interest rate minus servicing fee.
\item BorrowerRate: The borrower's interest rate for the loan.
\item Investors: The number of investors that funded the loan.
\item *.sc: The Variable * was scaled to have 0 mean and unit variance. (found only in regression results)
\end{itemize}

\subsection{Measuring Loan Performance with ROI}
<<roivstatus,cache=FALSE,fig.height=3,fig.cap="This chart shows the distribution of ROI for completed and defaulted loans over all three regimes. A continuous ROI variable offers more information about the quality of an investment than a binary default indicator. Dotted lines indicate mean ROI for category.",fig.height=3,fig.pos='H'>>=
nd$d.factor =  ifelse(nd$defaulted == 1, 'Defaulted','Completed')
smeans = summarise(group_by(nd,d.factor), mRoi = mean(roi, na.rm=TRUE))
ggplot(nd, aes(x=roi,fill=d.factor)) +
  geom_histogram(binwidth=.01)+ 
  geom_vline(data=smeans, aes(xintercept=mRoi,color=d.factor),linetype='longdash') + 
  theme_bw() +
  labs(title='Distribution of ROI for each Loan Outcome (all regimes)', x='ROI',y='Count',fill='Loan Status') +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-1,1,by=.25),label=percent) +
  facet_wrap(~d.factor,ncol=1,scales='free_y')
  @

@@

More generally, the slope of the curves in Figure \ref{fig:ratevscore} show that
late auction regime investors were much more sensitive to changes in borrower
credit score than early auction regime investors were. In the early auction
regime, a 100 point increase in credit score was associated with a 4\% reduction
in interest rate.\footnote{This analysis uses regressions that predict Interest
rate, not the chart itself} In the late auction regime, a 100 point increase in 
credit score is associated with an 11\% reduction. In the Prosper-rate regime,
a 100 point increase in credit score is associated with a 10\% reduction in interest rate.

@@

Prosper's reasoning for the change was to protect investors: by March 1, 2007, 14.9\% of loans to low-credit borrowers had already defaulted, more than triple the 4.6\% of loans to other borrowers that had already defaulted. \footnote{The 10\% gap in default rates is even more striking when the average age of the two groups of loans is considered. On March 1, 2007, the loans to low-credit borrowers averaged 100 days since origination, while loans to other borrowers averaged 122 days since origination. 14.9\% of low credit borrowers defaulted over an average time since origination of 100 days, whereas only 4.6\% of other borrowers defaulted over an average time since origination of 122 days.} Theoretically, however, it is more difficult to understand why Prosper needed to step in. From the EMH perspective, investors would have reacted to the news of the defaults on low credit loans by decreasing their exposure to low credit loans, or increasing the interest rates they bid on them. However, the median interest rate on low-credit, actually decreased from 24.23\% for the month of July, 2006 to 23.4\% for February, 2007, the month before the intervention. Similarly, exposure to low credit loans did not substantially decrease. On December 1, 2006, 13.3\% of these bad loans had already, after an average of only 3 monthly payments, gone into default, as compared to 3.7\% of other loans. Yet over the month of December 2006, loans to borrowers with credit score below 520 accounted for 22\% of orginations, the highest proportion in the site's history.  Hayek's wisdom of the crowd theory expects market participants to adjust their beliefs to new information more rapidly and flexibly than Prosper. The data, however, show that investors continued to make avoidably bad decisions until Prosper disallowed them.

@@

As the quantity of publicly available information increases, crowds will utilize an increasingly small fraction of it, leading either to inefficient prices or, perhaps, a new trend of "re-intermediation."
\newpage
\section{Appendix}
\subsection{Prosper's Charged-Off vs. Default Terminology}
After the shutdown, Prosper began using the term ``Charged-Off" to mean that a loan is delinquent past 120 days. We observe more defaults than charge-offs before the shutdown, but more charge-offs than defaults after, likely because of the shift in terminology. According to Prosper's website:
\begin{quote}
In general, a debt or account is considered “charged off” when it is unlikely that further payments will be received. Debts are usually charged off after they remain unpaid for a period of time (e.g., 90 to 180 days). Prosper uses the 120 days as the charge off threshold because loans that become over 120 days past due are eligible for sale to a debt buyer, and we have found that there is a steep drop-off in likelihood of further payments after 120 days of delinquency.\footcite{chargeoff}
\end{quote}
\subsection{Out of Sample Performance by Credit Grade}
<<retTab2,cache=FALSE>>=
vmeans = res %>% group_by(regime,model,grade) %>%  
  summarise(mn = sum((1+meanUL) *length) / sum(length) -1, n=sum(length))

@@

texreg(list(lm1.sc,lm2.sc,lm3.sc),custom.note='Y = Known-Default', label="kd2",
       ,custom.model.names = c('Auction-Early','Auction-Late','Prosper-rate')
       , caption='Coefficients extracted from Least Squares Regression to predict Known-Default in early auction, late auction, and Prosper-rate regime training data. Whereas the previous specification weights observations by maturity, this specification does not include weights, as a robustness check.',
       scalebox=.8, digits=4, bold=.05)
@
\newpage
\printbibliography
\end{document}

>>>>

\title{YALE DEPARTMENT OF ECONOMICS \\ \vspace*{60pt} The Folly of the Crowd: \\
Market Inefficiencies in Peer-to-Peer Lending}
\author{Samuel Shleifer
  \footnote{I am grateful to Professor Vineet Kumar for his help
  in developing the ideas and statistical methodologies involved in this paper,
  Devin Pope and Rajkamal Iyer for help manipulating the Prosper data, 
  and to Andrei Shleifer, Nicholas Barberis and Andrew Metrick for helpful comments and suggestions.}
  \footnote{R code available at \url{github.com/sshleifer/ProsperThesis}}\\
Thesis Advisor: Vineet Kumar\footnote{Yale School of Management}}
\maketitle

@@

\afterpage{\blankpage}
\newpage
\begin{abstract}
This essay analyzes the effect of market maker interventions on financial market
efficiency, using data from Prosper.com, the first p2p lending
website in the US.\footnote{P2P lending is the practice of individual and
  institutional investors lending money to borrowers over the internet, without
the presence of a financial intermediary} Between 2006 and 2011, Prosper underwent two
major policy changes that guided and then forced willing investors to lend at higher
interest rates, as well as to reduce their exposure to subprime borrowers.

@@

Freedman and Jin (2008) use investor
learning to explain this improvement. This paper expands on their work, 
with the advantage of three additional years of data and two major policy changes. 
I do not find much evidence for investor learning. Rather, I find evidence that
Prosper's rule changes guided and then forced willing investors to make better
decisions. When investors set interest rates by auction, investment strategies using publicly available
information earn excess returns, an inconsistency with the predictions of
the Efficient Markets Hypothesis. Once Prosper begins setting rates,
it is no longer possible to beat the market.
\end{abstract}
\newpage
\tableofcontents

@@

In the P2P Lending context, the market's price is the interest rate charged on a
loan. This paper studies the question of whether interest rates on Prosper.com
are set efficiently. Fama (1970)'s semi-strong EMH predicts that if interest rates were set efficiently, then...
\begin{enumerate}
\item Loans would have positive expected returns. Investors would be unwilling
  to lend to Borrowers if they expected to lose money. \footnote{Assuming a
  Risk-Free Rate of 0.}
\item Riskier loans would have higher expected returns. \footcite{sharpe}
\item It is not possible to do better than the average investor using publicly available information. \footcite{fama}
\end{enumerate}
Hypothesis 1 is tested by measuring historical returns of loans to borrowers of
each credit grade.
Hypothesis 2 is tested by comparing the returns of high risk loans to low risk
loans.
Hypothesis 3 is tested by using publicly available information to select loans
with a higher than average expected return, without look-ahead bias.

Prosper.com data provide some additional valuable insights into investor
behavior. Over the period I study, there are three interest rate determination regimes:
\begin{enumerate}
\item Early Auction Regime (April 19, 2006-October 15, 2008): Rates are set by
  investor auction. Investors have easy access to useful data, but Prosper does not guide them meaningfully.
\item Late Auction Regime: (July 20, 2009-December 19, 2010): Rates are set by
  investor auction, with Prosper offering valuable guidance in the form of more
  complex credit grading. 
\item Prosper-rate Regime: (December 20, 2010 - August 3, 2011): Prosper sets
  rates without investor input using its more complex credit grading algorithm.
\end{enumerate}
Additionally, for regimes (2) and (3), subprime borrowers are not allowed to
apply for Prosper loans.

@@

\begin{table}
  %\renewcommand\arraystretch{1.4}\arrayrulecolor{LightSteelBlue3}
  %\captionsetup{singlelinecheck=false, font=blue, labelfont=sc, labelsep=quad}
  \caption{Timeline of Events}\vskip -1.5ex
  \begin{tabular}{@{\,}r <{\hskip 2pt} !{\foo}
    >{\raggedright\arraybackslash}p{12cm}}
    \toprule
    \addlinespace[1.5ex]
    Apr. 21, 2006 & First loan of early auction regime.\\
    Mar. 1, 2007 & Prosper bans borrowers with credit score $\leq$ 520.\\
    Oct. 15, 2008 & Early Auction period ends with SEC enforced shutdown.\\
    Jul. 20, 2009 & Prosper reopens with more informative credit grades, requires repeat borrowers to have credit score
    $\geq$ 600, and requires new borrowers to have credit score $\geq$ 640.
    Beginning of Early Auction Regime.\\
    Dec. 19, 2010 & Prosper-rate regime, where Prosper sets rates directly,
    begins, because ``auctions made the deployment of funds more time consuming
    with little gain in lender returns.” - Prosper CEO Chris
    Larsen.\footcite{renton}\\
    Aug. 3, 2011 & Last loan in Prosper-rate regime originated.
  \end{tabular}
\label{timeline}
\end{table}

@@

As an illustrative example, suppose Bill the borrower needed to borrow \$100, and set his maximum rate at 20\%.
Let Chris, David and Ellen be investors, with Chris bidding (\$50, 15\%), David bidding (\$50, 18\%) and Ellen bidding (\$30,19\%). In this scenario, Chris and David would win the auction, and each lend lend Bill \$50, and receive an 18\% interest rate. Bill would pay 18\% interest before fees.
\subsection{Early Auction Regime}
In the early auction regime, loans originated between April 19, 2006 and October
15, 2008, auction participants had access to considerable information from
Prosper, including over 60  Borrower characteristics, such as residence,
employment status and an Experian credit score bucket, but few intelligent risk
estimates. Additionally, Prosper did not regulate who could apply for a loan on
the website. Besides basic fraud prevention, any applicant who received 100\%
funding from investors was eligible, for Prosper's first year of operation. On
March 1, 2007, amidst poor returns, Prosper announced that listings with credit
scores below 520 would no longer be offered to investors. This intervention is
not discussed in depth, for lack of space, but shows that after a year, more or less, Prosper felt the need to nudge a fully decentralized investment process in the right direction.

@@

Fama, (1970) notes the dependence of most tests of market efficiency on a model of risk and return.
The models, including most famously, Sharpe's capital asset pricing model (CAPM), make two predictions:
First,  since an investor can always hold cash risklessly, the average return on risky investments must be positive. Second, securities with higher systematic risk earn higher average returns.\footcite{sharpe}

@@

\subsection{Soft Information and Credit Screening}
The novel idea of P2P is a large group of investors may be better equipped to assess risk than traditional financial institutions. ``Traditionally, the credit score provided by rating agencies has been the main tool banks use to screen smaller borrowers."\footcite{iyer} As such, the initial strand of literature focuses largely on the novel sources of information utilized by P2P investors.
Gao and Lin (2012) analyze the text written by the borrower in the ``description
field" of the loan application, and find that descriptions that are easy to read have
2.3\% lower default rates than average, while complex narratives are associated
with a 3.6\% higher default rate. They do not measure returns.\footcite{gao}
Pope and Sydnor (2008) find that investors reduce
their defaults by statistically discriminating against racial minorities, but
also do not analyze return on investment.
Iyer et. al (2009) find that within the credit score buckets provided by Prosper, investors use soft information to further sort borrowers within buckets, reducing defaults.\footcite{iyer}

@@

\section{Data}
My sample consists of 39,555 three year, unsecured (no collateral) Loans, worth
\$231.7 Million,  originated from 2006-2011. The data was downloaded off Prosper.com.
Borrowers apply for loans by creating a public listing on the Prosper.com website, and can request anywhere from \$1,000 to \$35,000.\footnote{The minimum increased to \$2,000 effective Nov. 1, 2011}.
Since our analysis requires measurement of loan performance, which can only take
place three  years after origination (given 3 year maturity on all loans), the
sample is ended in 2011, as we do not know what will end up happening to more recently originated loans. 
As shown in Table \ref{sample}, this leaves us with many more loans from the
early auction regime than the post shutdown regimes.
<<sampTab, cache=FALSE>>=
samp.tab = mf(count(nd,regime))
colnames(samp.tab) = c('Regime','n','Freq.')
xtable(samp.tab,kdigits=4, caption='Observations per regime. Each observation is
one funded loan.', label='sample')
@

To help investors better assess the creditworthiness of borrowers, for every
loan originated in the site's history, Prosper posts historical data regarding loan
performance, such as payments made, interest rate, and principal lost.
Additionally, borrower
characteristics, including credit variables obtained from Experian and
verification of which are usually obtained directly
from Experian and verification of income and employment, are available to
investors for each listing. Moreover, historical market data is available for
download, through a query tool, and through third party websites. 
Interviews conducted at the 2008 Prosper Days Conference suggest that there was
``enormous heterogeneity in lender awareness of the data, ability to process the
data, and intent to track the data over time.''\footcite{freedman}

\subsection{Variable Definitions for the Prosper Data}
<<sumTable,cache=FALSE>>=
nd$default.rate =  nd$defaulted
k4 = c('roi','default.rate', 'LenderYield','BorrowerRate','regime','LoanOriginalAmount','Investors','CreditScoreEst','DebtToIncomeRatio', 'InquiriesLast6Months')
ag4 <- aggregate(.~regime, nd[k4], function(x) c(mean = mean(x), sd = sd(x)))
d4 = t(do.call(data.frame, c(ag4, check.names = FALSE)))
d5 = data.frame(d4)
names(d5) = d4[1,]
d6 = d5[-1,]
xtable(d6,caption='This table presents means (.mean) and standard deviations
(.sd) for relevant variables by regime.',digits=3, label='sumStats')
@
For each loan, Prosper offers 63 variables, I now define an important subset.
\begin{itemize}
\item BorrowerRate: Annual interest rate paid by the borrower.
\item TotalCreditLinespast7years: Number of credit lines, including cards and loans, used by the borrower in the seven years before the prosper loan was originated.
\item Inquirieslast6Months: Number of times a lender inquired about the Borrower's credit report, in the 6 months before the loan was originated. Traditional lenders make inquiries when a borrower applies for a loan, so this serves as a proxy for the number of times someone has borrowed in the past 6 months.
\item CurrentDelinquencies: Number of loans on which the borrower is delinquent at the origination date.
\item DelinquenciesLast7Years: The number of loans which the borrower has not repaid fully over the 7 years before the origination date.
\item MonthlyLoanPayment: The borrower's scheduled monthly payment on the Prosper loan.
\item CreditScoreRangeLower: The lower bound of the credit score range (which
  bounds his raw credit score) , that bounds a borrower's credit grade, as
  provided by Experian. Prosper provides the range instead of the raw score to
  preserve borrower anonymity.
\item CreditScoreEst: The average of the lower and upper bounds of the credit score range. 
\item PublicRecordsLast10Years: Number of public records of borrowing activity in the past 10 years at the origination date. 
\item DebtToIncomeRatio: The debt to income ratio of the borrower at the origination date. This value is capped at 10.01 (any debt to income ratio larger than 1000\% will be coded as 1001\%).
\item TotalProsperPaymentsBilled: Number of on time payments the borrower made on Prosper loans at the time they created the listing. This value is 0 if the borrower had no prior loans.
\item LenderYield: The lender's yield on the loan. Equal to the interest rate minus servicing fee.
\item BorrowerRate: The borrower's interest rate for the loan.
\item Investors: The number of investors that funded the loan.
\item *.sc: The Variable * was scaled to have 0 mean and unit variance. (found only in regression results)
\end{itemize}

\subsection{Measuring Loan Performance with ROI}
<<roivstatus,cache=FALSE,fig.height=3,fig.cap="This chart shows the distribution of ROI for all completed and defaulted loans in the sample. A continuous ROI variable offers more information about the quality of an investment than a binary default indicator. Dotted lines indicate mean ROI for category.",fig.height=3,fig.pos='H'>>=
nd$d.factor =  ifelse(nd$defaulted == 1, 'Defaulted','Completed')
smeans = summarise(group_by(nd,d.factor), mRoi = mean(roi, na.rm=TRUE))
ggplot(nd, aes(x=roi,fill=d.factor)) +
  geom_histogram(binwidth=.01)+ 
  geom_vline(data=smeans, aes(xintercept=mRoi,color=d.factor),linetype='longdash') + 
  theme_bw() +
  labs(title='Distribution of ROI for each Loan Outcome (all regimes)', x='ROI',y='Count',fill='Loan Status') +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-1,1,by=.25),label=percent) +
  facet_wrap(~d.factor,ncol=1,scales='free_y')
  @

@@

More generally, the slope of the curves in Figure \ref{fig:ratevscore} show that
late auction regime investors were much more sensitive to changes in borrower
credit score than early auction regime investors were. In the early auction
regime, a 100 point increase in credit score was associated with a 4\% reduction
in interest rate.\footnote{This analysis uses regressions that predict Interest
rate, not the chart itself. The regressions can be found in the appendix.} In the late auction regime, a 100 point increase in 
credit score is associated with an 11\% reduction. In the Prosper-rate regime,
a 100 point increase in credit score is associated with a 10\% reduction in interest rate.

@@

Prosper's reasoning for the change was to protect investors: by March 1, 2007,
14.9\% of loans to low-credit borrowers had already defaulted, more than triple
the 4.6\% of loans to other borrowers that had already defaulted. \footnote{The
10\% gap in default rates is even more striking when the average age of the two
groups of loans is considered. On March 1, 2007, the loans to low-credit
borrowers averaged 100 days since origination, while loans to other borrowers
averaged 122 days since origination. 14.9\% of low credit borrowers defaulted
over an average time since origination of 100 days, whereas only 4.6\% of other
borrowers defaulted over an average time since origination of 122 days.}
Theoretically, however, it is more difficult to understand why Prosper needed to
step in. From the EMH perspective, investors would have reacted to the news of
the defaults on low credit loans by decreasing their exposure to low credit
loans, or increasing the interest rates they bid on them. However, the median
interest rate on low-credit, actually decreased from 24.23\% for the month of
July, 2006 to 23.4\% for February, 2007, the month before the intervention.
Similarly, exposure to low credit loans did not substantially decrease. On
December 1, 2006, 13.3\% of these bad loans had already, after an average of
only 3 monthly payments, gone into default, as compared to 3.7\% of other loans.
Yet over the month of December 2006, loans to borrowers with credit score below
520 accounted for 22\% of originations, the highest proportion in the site's history.  Hayek's wisdom of the crowd theory expects market participants to adjust their beliefs to new information more rapidly and flexibly than Prosper. The data, however, show that investors continued to make avoidably bad decisions until Prosper disallowed them.

@@

\newpage
\printbibliography
\newpage
\section{Appendix}
\subsection{Prosper's Charged-Off vs. Default Terminology}
After the shutdown, Prosper began using the term ``Charged-Off" to mean that a loan is delinquent past 120 days. We observe more defaults than charge-offs before the shutdown, but more charge-offs than defaults after, likely because of the shift in terminology. According to Prosper's website:
\begin{quote}
In general, a debt or account is considered “charged off” when it is unlikely that further payments will be received. Debts are usually charged off after they remain unpaid for a period of time (e.g., 90 to 180 days). Prosper uses the 120 days as the charge off threshold because loans that become over 120 days past due are eligible for sale to a debt buyer, and we have found that there is a steep drop-off in likelihood of further payments after 120 days of delinquency.\footcite{chargeoff}
\end{quote}
\subsection{Out of Sample Performance by Credit Grade}
<<retTab2,cache=FALSE>>=
vmeans = res %>% group_by(regime,model,grade) %>%  
  summarise(mn = sum((1+meanUL) *length) / sum(length) -1, n=sum(length))

@@

texreg(list(lm1.sc,lm2.sc,lm3.sc),custom.note='Y = Known-Default', label="kd2",
       ,custom.model.names = c('Auction-Early','Auction-Late','Prosper-rate')
       , caption='Coefficients extracted from Least Squares Regression to predict Known-Default in early auction, late auction, and Prosper-rate regime training data. Whereas the previous specification weights observations by maturity, this specification does not include weights, as a robustness check.',
       scalebox=.8, digits=4, bold=.05)
@
\end{document}
